EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-14 04_cbinding.tex
Some more fixes

The c-bindings approach is extremely handy, in that is opens the way for interoperability 
with a large number of languages and frameworks. As an example, an experimental Julia API 
is provided in the EDIpack github repository. Rather than a fully functioning Julia package, 
this is at present a proof of concept, capable of replicating the results of Fortran, C++ and
python implementations for the Bethe lattice example driver.
Its structure and operation mimicks that of the EDIpy layer, with minimal language-specific differences.

The EDIpack Julia api consists of a module called {\tt EDIjl2}, which provides access to the global variables and functions contained in {\tt  libedipack2\_cbinding.so/.dylib}.

@@

\begin{itemize}
\item {\tt EDIPACK\_PATH}: if this environment variable is set, {\tt EDIjl} will look there for the library first
\item {\tt LD\_LIBRARY\_PATH}: for Linux systems
\item {\tt DYLD\_LIBRARY\_PATH}: for Mac systems
\end{itemize}

in partial analogy to the Python API, the global variable, but not the functions, are contained in a  {\tt struct} called {\tt global\_env}.

Similarly to the C++ case, global variables are loaded as raw pointers. As a consequence, it is important that the dimensions of the array variables (such as {\tt ULOC}) and, in general, the size of the memory sector occupied by each variable are correcty accounted for. This is achieved within the Julia module by appropriately casting the variables to compatible types, such as {\tt  Cint, Cbool, Cdouble}.
Functions are directly called from the dynamic library, making use of the {\tt  ccall} procedure. As previously stated, as a consequence of the C linking conventions multiple calls to {\tt  ccall} are necessary to account for the different input options of each interfaced Fortran procedure; the Julia wrapper functions take care of selecting the appropriate Fortran procedure depending on the set of input parameters provided by the user.

\subsubsection{Installation as usage}

@@

The correct way to access global variables and functions is then as in the following example:

@@

Particular care has to be taken with respect to parallelism: Julia offers an in-house platform for parallel computing with memory shearing through multi-threading. At present, there is no integration between this architecture and the MPI-based parallelism of \NAME{}, and their concurrent use in a single program is entirely untested and not encouraged. If the library is compiled with enabled parallelism, MPI support will need to be enabled in the user-created Julia program, by making use of modules such as {\tt  MPI.jl}. The {\tt MPI\_COMM\_WORLD} communicator will need to be correctly initialized by the user as well, and the program will need to be run with {\tt  mpiexec/mpirun}.

We provide an example script for the simple case of the Bethe lattice in the normal phase in the {\tt  examples} folder of the EDIjl2.0 repository. Documentation of the API and further examples are being developed.
\end{document}

>>>>

The c-bindings approach is extremely handy, in that is opens the way for interoperability 
with a large number of languages and frameworks. As an example, an experimental Julia API 
is provided in the EDIpack github repository. Rather than a fully functioning Julia package, 
this is at present a proof of concept, capable of replicating the results of Fortran, C++ and
python implementations for the Bethe lattice example driver.
Its structure and operation mimicks that of the EDIpy2 layer, with minimal language-specific differences.

The \NAME{} Julia api consists of a module called {\tt EDIjl2}, which provides access to the global variables and functions contained in {\tt  libedipack2\_cbinding.so/.dylib}.

@@

\begin{itemize}
\item {\tt EDIPACK\_PATH}: if this environment variable is set, {\tt EDIjl2} will look there for the library first
\item {\tt LD\_LIBRARY\_PATH}: for Linux systems
\item {\tt DYLD\_LIBRARY\_PATH}: for Mac systems
\end{itemize}

\subsubsection{Structure}

In partial analogy to the Python API, the global variables, but not the functions, are contained in a  {\tt struct} called {\tt global\_env}.

Similarly to the C++ case, global variables and functions are loaded as raw pointers. As a consequence, it is important that the dimensions of the array-like variables (such as {\tt ULOC}) and, in general, the amount of memory occupied by each variable are correcty accounted for. This is achieved within the Julia module by appropriately casting the variables to compatible types, such as {\tt  Cint, Cbool, Cdouble}.
Functions are called from the dynamic library, making use of the {\tt  ccall} procedure. As previously stated, as a consequence of the C linking conventions multiple alternative version of the interfaced Fortran procedures are present, to account for the different input variable combinations; the EDIjl2 wrapper functions take care of selecting the appropriate Fortran procedure depending on the set of input parameters provided by the user.

\subsubsection{Installation and usage}

@@

The correct way to access global variables and functions is as in the following example:

@@

Particular care has to be taken with respect to parallelism: Julia offers an in-house platform for parallel computing with memory sharing through multi-threading. At present, there is no integration between this architecture and the MPI-based parallelism of \NAME{}, and their concurrent use in a single program is entirely untested and not encouraged. If the \NAME{} library is compiled with enabled parallelism, MPI support will need to be enabled in the user-created Julia program, by making use of modules such as {\tt  MPI.jl}. The {\tt MPI\_COMM\_WORLD} communicator will need to be correctly initialized by the user as well, and the program will need to be run with {\tt  mpiexec/mpirun}.

We provide an example script for the simple case of the Bethe lattice in the normal phase in the {\tt  examples} folder of the EDIjl2.0 repository. This script is intended to be run serially. Documentation of the API and further examples are being developed.
\end{document}
